---
title: "Quiz2"
author: "Roberto"
date: "2/13/2020"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE, results = 'asis'}
library(kableExtra)
x = c(22,26,27,33,29,29,34,30,40)
y = c(497,541,556,576,578,607,662,739,805)
df = data.frame(x,y)
kable(df) %>%
  kable_styling(full_width = F, position = 'float_left', font_size = 7)
```
**a.**
**Look at a scatterplot of the data. Does it appear that a simple linear model is appropriate?** 

Yes.

We can also check the correlation, to check if the two variables are linearly associated.

Cor: 
```{r echo = FALSE}
cor(x,y)
```

```{r echo=FALSE, fig.width = 4, fig.height=4}
q1  = plot(x,y, xlab = ' x', ylab = 'y', pch = 19) 
abline(lm(y~x), col = 'red')
```


    **b.**
    **Which variable, X or Y, is the "response" variable        in this problem? Which is the "explanatory" or            "predictor"    variable?**
 
 
    The response variable is the dependent variable, which     is  *Y* in this problem. *X* is the predictor variable     or explanatory variable. *X* is the independent           variable.
 
    **c.**
    **What is the equation of the regression line? (Round       the slope and y-intercept to one decimal place.)**
 
    The estimated regression line equation is:
    ***y = 141.08 + 15.89x***

```{r echo = FALSE, fig.width=5, fig.height=3}
model = lm(formula = y ~ x)
model
summary(model)
```
 

    **d.**
    **Use the regression line to predict the value of Y         when   X = 31.**

```{r echo = FALSE }
Y = 141.08 + 15.89 * (31)
Y
```

